---
title: "Final Project"
author: "Shalom Mhanda"
format: pdf
editor: visual
---

```{r}
library(tidyverse) # for ggplot functions/plotting
library(broom) # for tidy() function
library(openintro) # for loan data set
library(GGally) # for ggpairs() function
```

# Research Question

Is number of bedrooms a good predictor of house price?

$H_O: \beta_1 = 0$

$H_A: \beta_1 ≠ 0$

```{r}
# Importing data
saratoga_ny_home_prices <- read.csv("/Users/smhanda/Documents/SDS220/Project/saratoga_ny_home_prices.csv")
```

```{r}
# Plotting Home price vs number of bedrooms
ggplot(data = saratoga_ny_home_prices, 
       aes(x = Bedrooms, 
           y = Price)) +
  geom_point(position = "jitter",
             size = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Home Price vs Number of Bedrooms")
```

## Checking assumptions

1.  Independence
2.  Linearity

```{r}
# Plotting linear regression line
ggplot(data = saratoga_ny_home_prices, 
       aes(x = Bedrooms, 
           y = Price)) +
  geom_point(position = "jitter",
             size = 0.5) +
  geom_smooth(se = F, method = "lm") +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Home Price vs Number of Bedrooms with Linear Regression Line")
```

3.  Normal residuals

```{r}
fit <- lm(Price ~Bedrooms, data = saratoga_ny_home_prices)
summary(fit)
```

```{r}
# Residual Plot
ggplot(fit, aes(x = .fitted, y = .resid)) +
  geom_point(position = "jitter",
             size = 0.5) +
  geom_hline(yintercept = 0)
```

```{r}
# Getting p value and confidence intervals
lm(Price ~ Bedrooms, data = saratoga_ny_home_prices) |>
tidy(conf.int = T, conf.level = .95)
```

Is the presence of a fireplace a good predictor of house price?

$H_O: \beta_1 = 0$

$H_A: \beta_1 ≠ 0$

```{r}
# Plotting Home price vs number of bedrooms
ggplot(data = saratoga_ny_home_prices, 
       aes(x = Fireplace, 
           y = Price,
           color = Fireplace)) +
  geom_point(position = "jitter",
             size = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Home Price vs Presence of Fireplace")
```
## Checking assumptions

1.  Independence
2.  Linearity

```{r}
# Plotting linear regression line
ggplot(data = saratoga_ny_home_prices, 
       aes(x = Fireplace, 
           y = Price)) +
  geom_point(position = "jitter",
             size = 0.5) +
  geom_smooth(se = F, method = "lm") +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Home Price vs Presence of Fireplace with Linear Regression Line")
```

3.  Normal residuals

```{r}
fit2 <- lm(Price ~Fireplace, data = saratoga_ny_home_prices)
summary(fit2)
```
```{r}
# Residual Plot
ggplot(fit2, aes(x = .fitted, y = .resid)) +
  geom_point(position = "jitter",
             size = 0.5) +
  geom_hline(yintercept = 0)
```

```{r}
# Getting p value and confidence intervals
lm(Price ~ Fireplace, data = saratoga_ny_home_prices) |>
tidy(conf.int = T, conf.level = .95)
```

Multiple Predictors: Number of Bedrooms and Presence of Fireplace
```{r}
ggplot(saratoga_ny_home_prices, aes(x = Bedrooms, y = Price, color = Fireplace)) +
  geom_point(position = "jitter",
             size = 0.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Home Price vs Presence of Fireplace and Number of Bedrooms")
```
```{r}
lm(Price ~ Bedrooms + Fireplace, data = saratoga_ny_home_prices) |>
tidy()
```

```{r}
# Checking R^2 adjusted
fit3 <- lm(Price ~ Bedrooms + Fireplace, data = saratoga_ny_home_prices)
summary(fit3)
```
## Check for Multicollinearity
```{r}
ggpairs(saratoga_ny_home_prices,
        columns = c("Price", "Bedrooms", "Fireplace"))
```

